---
title: "manova"
author: "Kar"
date: '2022-06-27'
output: html_document
---


# ONE WAY MANOVA

```{r}
library(tibble)
library(car)


```

```{r}
data("Baumann")

study <- Baumann %>% 
  dplyr::select(group, post.test.1, post.test.2, post.test.3)

```

*Assumption tests*

The first step is always test the assumption. 

1. Independent Observations Assumption

Look at variable 4, 5 and 6, they are post test results. 

```{r}
library(psych)

ICC(study[, 2:4])

```

From the Intraclass correlation coefficients, we only need to look at the first one "Single_raters_absolute", the type 1 ICC, the associated ICC is -0.456, it shows a **medium degree of correlation**. It is not too correlated that to the negative side (doesn't matter is positive or negative). 

So this assumption for the independent observations passess the test. 


*2. Normality test with mshapiro.test*

Normality test on individual continuous variables ONLY. 

```{r}

library(mvnormtest)

mshapiro.test(t(study[, -1]))

```

Ho is normality distributed, and so we cannot reject the null hypothesis and conclude that data are normally distributed. 


*3. Equal variance-covariance matrics assumption*


```{r}
library(biotools)

boxM(data = study[, -1], grouping = study[, 1])
  
```

We cannot reject the Ho that the levels in the grouping variable has the same variance-covariance matrics. 

*4. Determinant of the variance-covariance matrix.* 

```{r}
det(cov(study[, -1]))

```
Positive and pass the assumption. 


**One way MANOVA**

```{r}
oneway.manova <- manova(cbind(post.test.1, post.test.2, post.test.3) ~ group, data = study)
  
summary(oneway.manova, test = "Wilks")
summary(oneway.manova, test = "Pillai")
summary(oneway.manova, test = "Hotelling-Lawley")
summary(oneway.manova, test = "Roy")


```

All four test statistics have p-value of less than 0.05, we can reject the null hypothesis and conclude that the joint mean of 3 tests (post test 1, post test 2, and post test 3) among the 3 teaching techniques ARE the same. 


# Factorial MANOVA

3 Ho:

HO,1 : The joint mean of 3 variables (pH, Density and Conductivity) ARE NOT significantly different among the categories of the variable *Contours.*

HO,2 : The joint mean of 3 variables (pH, Density and Conductivity) ARE NOT significantly different among the categories of the variable *Depths*

HO,3 : The joint mean of the 3 variables (pH, Density, and Conductivity) ARE NOT statistically different for combinations of the Contour x Depths *interaction*

```{r}
library(car)
data(Soils)

Soils

```


```{r}
summary(Soils)

```


We will use pH, Dens, and Conductivity. 

```{r}
soils.lm <- lm(cbind(pH, Dens, Conduc) ~ Contour + Depth + Contour*Depth -1, 
               data = Soils)   # -1 remove intercept
  
summary(soils.lm)

```

* For all of the contour levels, p-values are lower than 0.05, so the contour variables is significantly affects the 3 variables (pH, Dens, Conduc). For the 3 variables, their means are different among the levels of contour.

* Also, the mean of the 3 outcome variables are different among the 3 levels of soil depth. 

* Interactions is considered not significant in influnencing the values of the 3 outcome variables.

Therefore, among 3 types of sum of squares:

* Type I SS --> For balanced data
* Type II SS --> Most powerful when there is no significant interactions terms
* Type III SS --> for when there is a significant interaction term

I can just use the type II SS in order to test the null hypothesis.




```{r}
Manova(soils.lm, multivariate = T, type = "II", test = "Wilks")
Manova(soils.lm, multivariate = T, type = "II", test = "Pillai")
Manova(soils.lm, multivariate = T, type = "II", test = "Hotelling-Lawley")
Manova(soils.lm, multivariate = T, type = "II", test = "Roy")
```

All interaction terms are insignificant. 

Conclusion:

* The joint mean of 3 variables (pH, Density, and Conductivity) are significantly different among the categories of the variables "Contour" and "Depths", p-values < 0.05.

* The joint mean of 3  variables (pH, Density, and Conductivity) are Not significantly different for combinations of the Contours*Depth interaction. 


 
